% !TeX spellcheck = en_GB
% Die erste (unkommentierte) Zeile im Dokument legt immer die
% Dokumentklasse fest
\documentclass{scrartcl} 

% Präambel:
% Einbinen von zusätzlichen Paketen. Falls für eine Datei keine Endung
% explizit angegeben wird, benutzt LaTeX '.tex'. Im Folgenden wird
% also die Datei 'edv_pakete.tex' eingebunden.
\input{edv_pakete}


% Verzeichnisse mit Abbildungen; kann gestrichen werden,
% falls Sie dies schon in edv_pakete.tex definiert haben:
%\graphicspath{{../report}}

\addbibresource{refs.bib} %Hinzufügen einer Literaturdatenbank aus dem angegebenen Verzeichnis

% Titel, Autor und Datum
\title{Computational Physics}
\subtitle{Exercise 2}
\date{\today}
\author{Christiane Groß, Nico Dichter}

% Jetzt startet das eigentliche Dokument
\begin{document}
	\maketitle
\section{The 2D-Ising-model}
Hamiltonian, expected values for magnetization and epsilon

\section{Deliberations}
\subsection{Numerical cost of calculation}
\paragraph{Energy} Have to iterate over all lattice points: effort proportional lambda.

\paragraph{change in energy for one spin flip}
naive: calculate two Hamiltonians, subtract: scales with lambda. 
Possible: only use nearest neighbours, as shown in lecture for 1D: effort only depends ion dimension, not lattice size.

\subsection{Meaning of J} critical J: phase transition from randomly ordered to ordered.

\subsection{determining errors}

\section{simulation}
In the implementation of the simulation gsl\_matrix objects get used to save the state of the largest lattice and views on these get used to create smaller lattices (see \cite{gsldoc_mat}). For each parameter set (N,h,J) a large amount of thermalization sweeps get made and afterwards several measurements of the wanted physical variable take place with single sweeps in between. A Sweep in this case means scanning through the complete lattice and performing a Metropolis-Hastings step at each site. The thermalization steps are needed to make sure the measurements take place on configurations which resemble the given parameter set according to the boltzmann distribution. To improve the run time the function energy\_change got implemented. It calculates $\Delta S$ by only using the parts of the Hamiltonian which change when $s_{i,j}$ changes to $-s_{i,j}$, these parts consist of the energy of $s_{i,j}$ in the external magnetic field and the coppling energy with the 4 neighbours.
So in total $\Delta S$ simplifies to:
\begin{equation}
	\Delta S=2 s_{i,j}\left(\dfrac{h}{T}+\dfrac{J}{T}(s_{i+1,j}+s_{i-1,j}+s_{i,j+1}+s_{i,j-1})\right)
\end{equation}


\section{Results}

\subsection{$\langle m\rangle$ for fixed $J$ and varying $h$}
plot, description

\subsection{$\langle \epsilon\rangle$ for fixed $h$ and varying $J$}
plot, description, comparison literature

\subsection{$\langle |m|\rangle$ for fixed $h$ and varying $J$}
plot, description, comparison literature
see sharp drop, phase transition
\paragraph{$\langle |m|\rangle$ vs. $\langle m\rangle$}
For $h=0$: everything cancels out, m=0 everywhere because simulation randomly flops between positive and negative behaviour, even at ordered phase due to finite length.
Use absolute value of m so it can't cancel out.

\paragraph{thermodynamical limit}



\newpage	
\listoffigures
\printbibliography
\end{document}